---
title: "Stat 437 HW1"
author: "Yu-Tung(Jenny), Cheng (11678647)"
header-includes:
- \usepackage{bbm}
- \usepackage{amssymb}
- \usepackage{amsmath}
- \usepackage{graphicx,float}
- \usepackage{natbib}
output:
  html_document:
    df_print: paged
  pdf_document: default
fontsize: 11pt
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# General rule

Please show your work and submit your computer codes in order to get points. Providing correct answers without supporting details does not receive full credits. This HW covers:

- The basics of `dplyr`
- Creating scatter plot using `ggplot2`
- Elementary Visualizations (via ggplot2): density plot, histogram, boxplot, barplot, pie chart
- Advanced Visualizations via ggplot2: faceting, annotation

For an assignment or project, you DO NOT have to submit your answers or reports using typesetting software. However, your answers must be well organized and well legible for grading. Please upload your answers in a document to the course space. Specifically, if you are not able to knit a .Rmd/.rmd file into an output file such as a .pdf, .doc, .docx or .html file that contains your codes, outputs from your codes, your interpretations on the outputs, and your answers in text (possibly with math expressions), please organize your codes, their outputs and your answers in a document in the format given below:

```
Problem or task or question ... 
Codes ...
Outputs ...
Your interpretations ...
```

It is absolutely not OK to just submit your codes only. This will result in a considerable loss of points on your assignments or projects. 

# Problem 1

Please refer to the NYC flight data `nycflights13` that has been discussed in the lecture notes and whose manual can be found at https://cran.r-project.org/web/packages/nycflights13/index.html. We will use `flights`, a tibble from `nycflights13`.


You are interested in looking into the average `arr_delay` for 6 different `month` 12, 1, 2, 6, 7 and 8, for 3 different `carrier` "UA", "AA" and "DL", and for `distance` that are greater than 700 miles, since you suspect that colder months and longer distances may result in longer average arrival delays. Note that you need to extract observations from `flights` and obtain the needed sample means for `arr_delay`, and that you are requird to use `dplyr` for this purpose.

The following tasks and questions are based on the extracted observations.

##install and set up R markdown
```{r, cache=TRUE}

library(knitr)
```

(1.a) In a single plot, create a density plot for `arr_delay` for each of the 6 months with `color` aesthetic designated by `month`. Note that you need to convert `month` into a factor in order to create the plot. What can you say about the average `arr_delay` across the 6 months?

```{r, cache=TRUE}

library(dplyr)
library(nycflights13)
library(ggplot2)
library(tidyverse)
head(flights)

#create a small data frame 'temp' from flights and select 6 months and 3 carriers from 'flights'
temp = flights %>% select(month, arr_delay, carrier, distance) %>% filter(month %in% c(12,1,2,6,7,8), carrier %in% c("UA","AA","DL"), distance > 700)

#remove rows having any NA
temp = na.omit(temp)

#obtain mean by each combination of month value and carrier
sm = temp %>% group_by(month, carrier) %>% summarise( mean_arr_delay = mean(arr_delay)) %>% as.data.frame()
head(sm)

#apply for the density plot with the select rows of arr_delay for each of the 6 months with color by month as factor
p2 = ggplot(sm, aes(x=mean_arr_delay, color = as.factor(month))) + geom_density(linetype = "dashed") + theme_bw()

print(p2)

  ##Interpretation: As the density plot with these 6 months as factors for computing the average arrival of delay, we can observe that February relatively stands the least frequency of arrival delay in 2013. The brown dashed line which is February with the positively skewed right distribution mostly gathered on the left which its mode is smaller than mean.  


```

(1.b) In a single plot, create a boxplot for `arr_delay` for each of the 3 carriers. What can you say about the average `arr_delay` for the 3 carriers?

```{r, cache=TRUE}

#create a boxplot

p3= ggplot(sm, aes(x=mean_arr_delay, y= carrier, fill = carrier)) + geom_boxplot() + theme_bw()+ stat_summary(fun.y=mean,geom="point",shape=23, size=4, col = "white")

print(p3)

 ##Interpretation: The median of AA carrier is relatively small on the left-most point which is the less frequency of arrival delay as the box plot with x-axis of  arrival delay. However, as the scatter degree with its median of AA is more concentrated from the left side which means its data distribution is left-skewed. For DL carrier, there is a huge range  between the upper quartile and the lower quartile compared with AA and UA, and it shows the arrival delay of DL is unstable. As for UA, the upper quartile, lower quartile, and the median are bigger than others since the arrival delay of UA is more than others.    
```

(1.c) Create a pie chart for the 3 carriers where the percentages are the proportions of observations for each carrier and where percentages are superimposed on the sectors of the pie chart disc.

```{r, cache=TRUE}

library(scales)
sm2 = temp %>% group_by(carrier) %>% count() %>% ungroup() %>% mutate(percentage=`n`/sum(`n`))  %>% arrange(desc(carrier))

sm2$labels <- scales::percent(sm2$percentage)

print(sm2)

pie = ggplot(sm2) + geom_bar(aes(x="", y =percentage, fill=carrier), stat="identity", width = 1)+coord_polar("y", start = 0) + theme_void() + geom_text(aes(x=1, y= cumsum(percentage)-percentage/2, label=labels))

print(pie)

##Interpretation: The pie chart expresses that UA has 41.7%, DL has 33.7%, and AA has 24.6% of their observations.

```

(1.d) Plot `arr_delay` against `distance` with `facet_grid` designated by `month` and `carrier`.

```{r, cache=TRUE}

plot1 = ggplot(data = temp) + geom_point(mapping = aes(x = arr_delay, y = distance))

plot2 = plot1 + facet_wrap(month~carrier, nrow = 4)

print(plot2)

##Interpretation: Only UA carrier has the records of arriving delay with the long-distance flights. 




```


(1.e) For each feasible combination of values of `month` and `carrier`, compute the sample average of `arr_delay` and save them into the variable `mean_arr_delay`, and compute the sample average of `distance` and save these averages into the variable `mean_distance`. 
 Plot `month` against `mean_arr_delay` with `shape` designated by `carrier` and `color` by `mean_distance` and annotate each point by its associated `carrier` name. 

```{r, cache=TRUE}

#obtain mean by each combination of month value and carrier
mean_arr_delay = temp %>% group_by(month, carrier) %>% summarise( mean_arr_delay = mean(arr_delay)) %>% as.data.frame()

print(mean_arr_delay)

mean_distance = temp %>% group_by(month, carrier) %>% summarise( mean_distance = mean(distance)) %>% as.data.frame()

print(mean_distance)

sm['mean_distance'] <- c(mean_distance)

print(sm)


#Plot `month` against `mean_arr_delay` with `shape` designated by `carrier` and `color` by `mean_distance` 
# and annotate each point by its associated `carrier` name
pp3 = ggplot(sm, aes(x=as.factor(month), y = mean_arr_delay)) + geom_point(aes(as.factor(month), mean_arr_delay, colour = mean_distance, shape = carrier)) + geom_text(data=sm, aes(x=as.factor(month) , y=mean_arr_delay, label=carrier)) + scale_fill_distiller(palette = "YlOrBr") + scale_shape_manual(values = 1:length(unique(sm$carrier)) )

print(pp3)
##Interpretation: DL is unstable and it delays pretty often in July.


```

# Problem 2

Please refer to the data set `mpg` that is available from the `ggplot2` package. Plot `displ` against `hwy` with faceting by `drv` and `cyl`, `color` disgnated by `class`, and `shape` by `trans`. This illustrates visualization with 4 factors.

```{r, cache=TRUE}
#visualization with >= 3 factors

p1c = ggplot(mpg, aes(x = displ, y = hwy)) + theme_bw() + geom_point(aes(color=class, shape=trans)) + facet_grid(drv~cyl) + scale_shape_manual(values=c(1:10))

print(p1c)

##Interpretation: Larger cars has more cylinders and higher engine displacement in litres, and the smaller cars requires less.

```



